---
title: "End-to-End: From Evidence to Submission Decision"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{End-to-End: From Evidence to Submission Decision}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.width = 7,
  fig.height = 4
)
```

This vignette walks through the complete R4SUB workflow from loading evidence
to obtaining a regulatory-calibrated Submission Confidence Index (SCI).

## 1. Load the ecosystem

One call loads all R4SUB packages:

```{r load}
library(r4sub)
```

## 2. Inspect available example data

```{r datasets}
library(r4subdata)
list_datasets()
```

## 3. Load evidence

`evidence_pharma` is a 250-row evidence table covering study CDISCPILOT01
across all four readiness pillars.

```{r evidence}
data(evidence_pharma)

# Schema overview
str(evidence_pharma[, c("indicator_domain", "severity", "result",
                         "asset_type", "indicator_id")])

# Pass / warn / fail breakdown
evidence_summary(evidence_pharma)
```

## 4. Compute the Submission Confidence Index

```{r score}
# Indicator-level scores
ind_scores <- compute_indicator_scores(evidence_pharma)
head(ind_scores)

# Pillar scores (quality, trace, risk, usability)
pillar_scores <- compute_pillar_scores(evidence_pharma)
pillar_scores

# SCI
sci <- compute_sci(pillar_scores)
cat("SCI:", sci$SCI, "\n")
cat("Band:", sci$band, "\n")
```

```{r pillar-chart, fig.cap="Pillar scores contributing to the SCI"}
scores_vec <- setNames(pillar_scores$score, pillar_scores$pillar)
cols <- c(quality   = "#2C6DB5",
          trace     = "#27AE60",
          risk      = "#E74C3C",
          usability = "#F39C12")

par(mar = c(4, 6, 3, 2))
barplot(
  scores_vec[names(cols)],
  horiz  = TRUE,
  las    = 1,
  col    = cols[names(scores_vec)],
  border = NA,
  xlim   = c(0, 1),
  xlab   = "Score (0–1)",
  main   = paste0("Pillar Scores  |  SCI = ", round(sci$SCI, 1),
                  "  [", sci$band, "]")
)
abline(v = 0.85, lty = 2, col = "#555555")
```

## 5. Understand what's driving the score

```{r explain}
explanation <- sci_explain(evidence_pharma)
head(explanation$top_loss, 8)
```

## 6. Apply a regulatory authority profile

Calibrate the SCI weights for an FDA NDA submission:

```{r profile}
prof <- submission_profile("FDA", "NDA")
prof$pillar_weights

# Validate evidence completeness against the profile
val <- validate_against_profile(evidence_pharma, prof)
cat("Compliant:", val$is_compliant, "\n")
cat("Coverage: ", round(val$coverage * 100, 1), "%\n", sep = "")

if (length(val$missing_indicators) > 0) {
  cat("Missing indicators:\n")
  print(val$missing_indicators)
}
```

```{r profile-sci}
# Re-score using FDA-specific pillar weights
pillar_fda <- compute_pillar_scores(evidence_pharma,
                                    config = profile_sci_config(prof))
sci_fda    <- compute_sci(pillar_fda, config = profile_sci_config(prof))
cat("FDA-calibrated SCI:", sci_fda$SCI, "[", sci_fda$band, "]\n")
```

## 7. Compare authority profiles

```{r compare-profiles, fig.cap="SCI under different regulatory authority profiles"}
authorities <- list(
  "FDA / NDA"  = submission_profile("FDA",  "NDA"),
  "EMA / MAA"  = submission_profile("EMA",  "MAA"),
  "PMDA / NDA" = submission_profile("PMDA", "NDA_JP")
)

sci_vals <- vapply(authorities, function(p) {
  compute_sci(
    compute_pillar_scores(evidence_pharma, config = profile_sci_config(p)),
    config = profile_sci_config(p)
  )$SCI
}, numeric(1))

par(mar = c(5, 8, 3, 2))
barplot(
  sci_vals,
  horiz  = TRUE,
  las    = 1,
  col    = "#2C6DB5",
  border = NA,
  xlim   = c(0, 100),
  xlab   = "SCI (0–100)",
  main   = "SCI by Regulatory Authority Profile"
)
abline(v = 85, lty = 2, col = "#27AE60")
abline(v = 70, lty = 2, col = "#F39C12")
text(86, 0.7, "Ready", col = "#27AE60", cex = 0.8, adj = 0)
text(71, 1.9, "Minor gaps", col = "#F39C12", cex = 0.8, adj = 0)
```

## 8. Risk assessment

```{r risk}
data(risk_register_pharma)
rr     <- create_risk_register(risk_register_pharma)
scores <- compute_risk_scores(rr)

cat("Mean RPN:    ", round(scores$mean_rpn, 1), "\n")
cat("Max RPN:     ", scores$max_rpn, "\n")
cat("Critical risks:", scores$n_critical, "\n")
cat("High risks:    ", scores$n_high, "\n")
```

```{r risk-chart, fig.cap="Risk distribution by severity level"}
dist <- scores$level_distribution
par(mar = c(4, 4, 3, 2))
barplot(
  dist,
  col    = c(low = "#27AE60", medium = "#F39C12",
             high = "#E67E22", critical = "#E74C3C")[names(dist)],
  border = NA,
  ylab   = "Count",
  main   = "Risk Register: Level Distribution",
  las    = 1
)
```

## 9. Traceability coverage

```{r trace}
data(adam_metadata)
data(sdtm_metadata)
data(trace_mapping)

tm  <- build_trace_model(adam_metadata, sdtm_metadata, trace_mapping)
ctx <- r4sub_run_context(study_id = "CDISCPILOT01", environment = "review")
ev_trace <- trace_model_to_evidence(tm, ctx = ctx,
                                    source_name    = "r4subtrace",
                                    source_version = "0.1.0")

ind_trace <- trace_indicator_scores(ev_trace)
ind_trace[, c("indicator_id", "indicator_name", "score", "result")]
```

## 10. Sensitivity analysis

How stable is the SCI if pillar weights shift?

```{r sensitivity, fig.cap="SCI sensitivity to pillar weight variations"}
sens <- sci_sensitivity_analysis(evidence_pharma)
scenarios <- sens$scenarios

par(mar = c(5, 10, 3, 2))
barplot(
  setNames(scenarios$sci, scenarios$scenario),
  horiz  = TRUE,
  las    = 1,
  col    = "#2C6DB5",
  border = NA,
  xlim   = c(0, 100),
  xlab   = "SCI (0–100)",
  main   = "SCI Sensitivity by Weight Scenario"
)
abline(v = sens$base_sci, lty = 2, col = "#E74C3C")
```

## 11. Launch the dashboard

The interactive Shiny dashboard provides a point-and-click interface for all
of the above:

```r
library(r4subui)
r4sub_app(evidence = evidence_pharma)
```

The dashboard opens in your browser with eight tabs:
**Overview · Evidence · Indicators · Pillars · Sensitivity · Risk · Traceability · Authority**

## Summary

```{r summary-table, echo = FALSE}
knitr::kable(
  data.frame(
    Step      = c("Load data", "Score", "Profile", "Risk", "Trace", "Dashboard"),
    Package   = c("r4subdata", "r4subscore", "r4subprofile",
                  "r4subrisk", "r4subtrace", "r4subui"),
    `Key output` = c(
      "evidence_pharma (250 rows, 4 pillars)",
      paste0("SCI = ", round(sci$SCI, 1), " [", sci$band, "]"),
      paste0("FDA NDA: ", round(sci_fda$SCI, 1), " | Compliant: ", val$is_compliant),
      paste0("Mean RPN = ", round(scores$mean_rpn, 1),
             ", Critical = ", scores$n_critical),
      paste0(nrow(ev_trace), " trace evidence rows"),
      "Interactive browser dashboard"
    ),
    check.names = FALSE
  ),
  align = "lll"
)
```
