---
title: "End-to-End: From Evidence to Submission Decision"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{End-to-End: From Evidence to Submission Decision}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.width = 7,
  fig.height = 4
)
```

This vignette walks through the complete R4SUB workflow — from loading evidence
to a regulatory-calibrated Submission Confidence Index (SCI) and decision band.

## 1. Load the ecosystem

```{r load}
library(r4sub)
library(r4subdata)
```

## 2. Load evidence

`evidence_pharma` contains 250 evidence rows for study CDISCPILOT01 across
all four readiness pillars.

```{r evidence}
data(evidence_pharma)

# Pass / warn / fail by pillar
evidence_summary(evidence_pharma)
```

## 3. Compute the Submission Confidence Index

```{r score}
pillar_scores <- compute_pillar_scores(evidence_pharma)
pillar_scores[, c("pillar", "pillar_score", "n_indicators", "weight")]

sci <- compute_sci(pillar_scores)
cat("SCI:", sci$SCI, "\n")
cat("Band:", sci$band, "\n")
```

```{r pillar-chart, fig.cap = "Pillar scores contributing to the SCI"}
scores_vec <- setNames(pillar_scores$pillar_score, pillar_scores$pillar)
cols <- c(quality = "#2C6DB5", trace = "#27AE60",
          risk = "#E74C3C", usability = "#F39C12")

par(mar = c(4, 6, 3, 2))
barplot(
  scores_vec[names(cols)],
  horiz  = TRUE,
  las    = 1,
  col    = cols[names(scores_vec)],
  border = NA,
  xlim   = c(0, 1),
  xlab   = "Score (0-1)",
  main   = paste0("Pillar Scores  |  SCI = ", round(sci$SCI, 1),
                  "  [", sci$band, "]")
)
abline(v = 0.85, lty = 2, col = "#555555")
```

## 4. Understand what is driving the score

```{r explain}
expl     <- sci_explain(evidence_pharma)
top_loss <- head(expl$indicator_contributions, 8)
top_loss[, c("indicator_id", "indicator_name", "indicator_score", "loss")]
```

```{r loss-chart, fig.cap = "Top 8 indicators by SCI loss contribution"}
par(mar = c(4, 14, 3, 2))
barplot(
  rev(top_loss$loss),
  names.arg = rev(top_loss$indicator_name),
  horiz  = TRUE,
  las    = 1,
  col    = "#E74C3C",
  border = NA,
  xlab   = "SCI loss",
  main   = "Top Loss Contributors"
)
```

## 5. Apply a regulatory authority profile

```{r profile}
prof <- submission_profile("FDA", "NDA")
prof$pillar_weights

val <- validate_against_profile(evidence_pharma, prof)
cat("Compliant:", val$is_compliant, "\n")
cat("Coverage: ", round(val$coverage * 100, 1), "%\n", sep = "")
```

```{r profile-sci}
cfg_fda    <- profile_sci_config(prof)
ps_fda     <- compute_pillar_scores(evidence_pharma, config = cfg_fda)
sci_fda    <- compute_sci(ps_fda, config = cfg_fda)
cat("FDA-calibrated SCI:", sci_fda$SCI, "[", sci_fda$band, "]\n")
```

## 6. Compare regulatory authority profiles

```{r compare-profiles, fig.cap = "SCI under different regulatory authority profiles"}
profiles <- list(
  "FDA / NDA"    = submission_profile("FDA",  "NDA"),
  "EMA / MAA"    = submission_profile("EMA",  "MAA"),
  "PMDA / NDA"   = submission_profile("PMDA", "NDA_JP")
)

sci_vals <- vapply(profiles, function(p) {
  cfg <- profile_sci_config(p)
  compute_sci(compute_pillar_scores(evidence_pharma, config = cfg),
              config = cfg)$SCI
}, numeric(1))

par(mar = c(4, 9, 3, 2))
barplot(
  sci_vals,
  horiz  = TRUE,
  las    = 1,
  col    = "#2C6DB5",
  border = NA,
  xlim   = c(0, 100),
  xlab   = "SCI (0-100)",
  main   = "SCI by Regulatory Authority Profile"
)
abline(v = 85, lty = 2, col = "#27AE60")
abline(v = 70, lty = 2, col = "#F39C12")
```

## 7. Risk assessment

```{r risk}
data(risk_register_pharma)
rr           <- create_risk_register(risk_register_pharma)
risk_scores  <- compute_risk_scores(rr)

cat("Mean RPN:        ", risk_scores$mean_rpn, "\n")
cat("Max RPN:         ", risk_scores$max_rpn, "\n")
cat("Total risks:     ", risk_scores$n_risks, "\n")

print(risk_scores$risk_distribution)
```

```{r risk-chart, fig.cap = "Risk distribution by severity level"}
rd        <- risk_scores$risk_distribution
risk_cols <- c(low = "#27AE60", medium = "#F39C12",
               high = "#E67E22", critical = "#E74C3C")
rd_named  <- setNames(rd$n, rd$risk_level)
rd_plot   <- rd_named[intersect(names(risk_cols), names(rd_named))]

par(mar = c(4, 5, 3, 2))
barplot(
  rd_plot,
  col    = risk_cols[names(rd_plot)],
  border = NA,
  ylab   = "Count",
  main   = paste0("Risk Distribution  |  Mean RPN = ", risk_scores$mean_rpn),
  las    = 1
)
```

## 8. Traceability coverage

```{r trace}
data(adam_metadata)
data(sdtm_metadata)
data(trace_mapping)

ctx      <- r4sub_run_context(study_id = "CDISCPILOT01", environment = "review")
tm       <- build_trace_model(adam_metadata, sdtm_metadata, trace_mapping)
ev_trace <- trace_model_to_evidence(tm, ctx = ctx)

ind_trace <- trace_indicator_scores(ev_trace)
ind_trace[, c("indicator_id", "indicator_name", "score", "result")]
```

## 9. Sensitivity analysis

```{r sensitivity, fig.cap = "SCI sensitivity to pillar weight variations"}
weight_grid <- data.frame(
  quality   = c(0.35, 0.50, 0.25, 0.25),
  trace     = c(0.25, 0.20, 0.40, 0.25),
  risk      = c(0.25, 0.20, 0.25, 0.25),
  usability = c(0.15, 0.10, 0.10, 0.25)
)
scenario_labels <- c("Default", "Quality-heavy", "Trace-heavy", "Equal")

sens <- sci_sensitivity_analysis(evidence_pharma, weight_grid)

par(mar = c(4, 11, 3, 2))
barplot(
  setNames(sens$SCI, scenario_labels),
  horiz  = TRUE,
  las    = 1,
  col    = "#2C6DB5",
  border = NA,
  xlim   = c(0, 100),
  xlab   = "SCI (0-100)",
  main   = "SCI Sensitivity by Weight Scenario"
)
abline(v = c(70, 85), lty = 2, col = "#888888")
```

## 10. Launch the dashboard

```r
library(r4subui)
r4sub_app(evidence = evidence_pharma)
```

The dashboard opens in your browser with eight tabs:
**Overview · Evidence · Indicators · Pillars · Sensitivity · Risk · Traceability · Authority**

## Summary

```{r summary-table, echo = FALSE}
n_critical <- sum(rr$risk_level == "critical", na.rm = TRUE)

knitr::kable(
  data.frame(
    Step    = c("Load data", "Score", "Profile", "Risk", "Trace", "Dashboard"),
    Package = c("r4subdata", "r4subscore", "r4subprofile",
                "r4subrisk", "r4subtrace", "r4subui"),
    Output  = c(
      "evidence_pharma: 250 rows, 4 pillars",
      paste0("SCI = ", round(sci$SCI, 1), " [", sci$band, "]"),
      paste0("FDA NDA: ", round(sci_fda$SCI, 1),
             " | Compliant: ", val$is_compliant),
      paste0("Mean RPN = ", risk_scores$mean_rpn,
             ", Critical = ", n_critical),
      paste0(nrow(ev_trace), " trace evidence rows"),
      "Interactive browser dashboard"
    ),
    check.names = FALSE
  ),
  align = "lll"
)
```
