---
title: "R4SUB and pharmaverse"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{R4SUB and pharmaverse}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

pharmaverse builds submissions. R4SUB scores whether they are ready to file.

The two fit together cleanly because they sit at different points in the same
pipeline. A team uses admiral to derive ADaM, metacore to hold the
specification, and xportr to write compliant transport files. R4SUB takes what
those tools produce and answers a separate question: is this package ready to go
to a regulator, and if not, what is holding it back.

The handoff between the two worlds is the R4SUB evidence table. Anything that can
describe a dataset or a check can write evidence rows, and everything downstream
scores from there. This article shows the three main integration points.

```{r libs, message = FALSE}
library(r4subscore)
library(r4subtrace)
library(r4subdata)
```

## admiral to R4SUB

admiral derives ADaM datasets. R4SUB reads the shape of those datasets and their
mapping back to SDTM to score traceability and usability.

A typical admiral step looks like this. It is shown for context and is not run
here, because it needs the admiral toolchain:

```{r admiral-derive, eval = FALSE}
library(admiral)
library(pharmaversesdtm)

adsl <- pharmaversesdtm::dm |>
  derive_vars_merged(
    dataset_add = pharmaversesdtm::ex,
    new_vars = exprs(TRTSDT = convert_dtc_to_dt(EXSTDTC)),
    by_vars = exprs(USUBJID)
  )
```

What R4SUB needs from the result is metadata: the variables, their labels, and
how they map to source. Once you have that, the readiness side is ordinary
R4SUB. The example metadata in `r4subdata` has exactly the shape admiral output
produces once you extract its variable list.

```{r admiral-score}
tm <- build_trace_model(adam_metadata, sdtm_metadata, mapping = trace_mapping)
ctx <- r4subcore::r4sub_run_context(study_id = "CDISCPILOT01",
                                    environment = "UAT")
trace_ev <- trace_model_to_evidence(tm, ctx = ctx)
trace_indicator_scores(trace_ev)
```

The admiral-derived datasets go in; traceability coverage comes out, ready to
feed the Trace pillar of the Submission Confidence Index. The `r4subpharma`
package provides adapters that go straight from metacore and admiral objects to
this evidence, so in practice you rarely assemble the metadata by hand.

## xportr to R4SUB

xportr applies SAS transport metadata to a data frame: variable labels, types,
lengths, and formats, taken from the specification. That same metadata is what
the Quality and Usability pillars read.

The xportr step, shown for context:

```{r xportr, eval = FALSE}
library(xportr)

adsl_xpt <- adsl |>
  xportr_type(spec) |>
  xportr_length(spec) |>
  xportr_label(spec) |>
  xportr_format(spec)
```

After xportr runs, the variable attributes are exactly the columns R4SUB checks:
`label`, `type`, `length`, and `format`. The example ADaM metadata carries those
already, so the usability read is a direct call.

```{r xportr-score, message = FALSE}
library(r4subusability)

# xportr output described as metadata: label, type, length, format
meta <- adam_metadata
meta$origin <- ifelse(is.na(meta$format), "Derived", "Predecessor")

usab_ev <- usability_indicators(meta, assets = c("define.xml"), ctx = ctx)
usability_summary(usab_ev)
```

Labels and types that xportr populated become passing usability checks. Anything
xportr could not fill from the spec shows up as a gap, which is precisely the
signal a reviewer would raise.

## riskmetric and r4subrisk

riskmetric and r4subrisk both have "risk" in the name, and they answer opposite
halves of the same worry.

riskmetric scores the R packages you build with: is admiral, or your internal
derivation package, mature and well tested enough to depend on? That is tooling
risk, and it is shown here for context:

```{r riskmetric, eval = FALSE}
library(riskmetric)

pkg_ref("admiral") |>
  pkg_assess() |>
  pkg_score()
```

r4subrisk scores the submission itself: given the datasets and findings, how
risky is this package to file? That is deliverable risk, computed with FMEA.

```{r r4subrisk-score, message = FALSE}
library(r4subrisk)

rr <- create_risk_register(risk_register_pharma)
compute_risk_scores(rr)$overall_risk_score
```

The two are complementary. Use riskmetric to decide the tools are trustworthy,
then use r4subrisk to score whether the thing they produced is ready. One vets
the factory, the other inspects the product.

## Where R4SUB fits

| Stage | pharmaverse | R4SUB |
|---|---|---|
| Build ADaM | admiral, metacore | reads the result |
| Write transport files | xportr | reads label, type, length, format |
| Vet the tooling | riskmetric | assumes trustworthy tools |
| Score the submission | not the goal | evidence, SCI, risk, traceability |

pharmaverse gets the datasets built and compliant. R4SUB takes it from there and
turns "we think it is ready" into a number you can defend.
